# Research Question 4 {#rq-4}

We have already partially answered this question in the previous section by comparing the statistics across each approach and metric. To partially answer the last research question, we primarily compare statistics of the __low-regret areas__ created for each climate-smart approach. As mentioned in the previous section, these areas are selected as climate refugia regardless of the climate metric used (for each approach).

Here are the low-regret areas for all approaches
```{r include = FALSE}
ggLowRegret_Feature <- plot_lowregret(LowRegret_Feature, land) + labs(subtitle = "Feature, SSP 5-8.5")
ggLowRegret_Percentile <- plot_lowregret(LowRegret_Percentile, land) + labs(subtitle = "Percentile, SSP 5-8.5")
ggLowRegret_Penalty <- plot_lowregret(LowRegret_Penalty, land) + labs(subtitle = "Penalty, SSP 5-8.5")
```
```{r}
(ggLowRegret_Feature + ggLowRegret_Percentile) / (ggLowRegret_Penalty + plot_spacer()) + plot_layout(guides = "collect")
```

The summaries for each of the low-regret areas
```{r include = FALSE}
LowRegret_SummaryAll <- LowRegret_SummaryFeature %>% 
  rbind(., LowRegret_SummaryPercentile) %>% 
  rbind(., LowRegret_SummaryPenalty)

# Calculate climate metrics for low-regret areas
LowRegretFeature_df <- LowRegret_Feature %>% 
  as_tibble() %>% 
  dplyr::select(-feature_velocity_585, -feature_o2os_585, -feature_phos_585, -feature_tos_585, -Cost, -Cost_squish) %>%
  dplyr::rename(SelectionFeature = selection) %>% 
  left_join(., roc_tos_SSP585 %>% as_tibble() %>% dplyr::select(slpTrends, geometry), by = "geometry") %>% 
  dplyr::rename(tos = slpTrends, Feature_Solution = solution_1) %>% 
  left_join(., roc_phos_SSP585 %>% as_tibble() %>% dplyr::select(slpTrends, geometry), by = "geometry") %>% 
  dplyr::rename(phos = slpTrends) %>% 
  left_join(., roc_o2os_SSP585 %>% as_tibble() %>% dplyr::select(slpTrends, geometry), by = "geometry") %>% 
  dplyr::rename(o2os = slpTrends) %>% 
  left_join(., velocity_SSP585 %>% as_tibble() %>% dplyr::select(voccMag, geometry), by = "geometry") %>% 
  dplyr::rename(velocity = voccMag)

LowRegretPenalty_df <- LowRegret_Penalty %>% 
  as_tibble() %>% 
  dplyr::select(-penalty_velocity_585, -penalty_o2os_585, -penalty_phos_585, -penalty_tos_585, -Cost, -Cost_squish) %>%
  dplyr::rename(SelectionPenalty = selection, Penalty_Solution = solution_1)

LowRegretPercentile_df <- LowRegret_Percentile %>% 
  as_tibble() %>% 
  dplyr::select(-percentile_velocity_585, -percentile_o2os_585, -percentile_phos_585, -percentile_tos_585, -Cost, -Cost_squish) %>% 
  dplyr::rename(SelectionPercentile = selection, Percentile_Solution = solution_1)

LowRegretAll_sf <- left_join(LowRegretFeature_df, LowRegretPercentile_df, by = c("cellID", "geometry")) %>% 
  left_join(., LowRegretPenalty_df, by = c("cellID", "geometry")) %>% 
  dplyr::mutate(SelectionAll = case_when((SelectionFeature == 4 & SelectionPercentile == 4 & SelectionPenalty == 4) ~ 1, TRUE ~ 0)) %>% 
  st_as_sf(sf_column_name = "geometry")

LowRegret_SummaryAll <- lowRegret_ClimateSummary(df = LowRegretAll_sf, approach_column = "Feature_Solution") %>% 
  rbind(., lowRegret_ClimateSummary(df = LowRegretAll_sf, approach_column = "Percentile_Solution")) %>% 
  rbind(., lowRegret_ClimateSummary(df = LowRegretAll_sf, approach_column = "Penalty_Solution")) %>% 
  as_tibble() %>% 
  left_join(., LowRegret_SummaryAll, by = "approach")
  

# Plots
ggComparison_Cost_LowRegret <- plot_LowRegretStatistics(LowRegret_SummaryAll, col_name = "log10(total_cost)", y_axis = "log10(cost)")
ggComparison_Area_LowRegret <- plot_LowRegretStatistics(LowRegret_SummaryAll, col_name = "percent_area", y_axis = "% area")

ggComparison_Warming_LowRegret <- plot_LowRegretStatistics(LowRegret_SummaryAll, col_name = "mean_climate_warming", y_axis = expression('Δ'^"o"*'C yr'^"-1"*''))
ggComparison_LogWarming_LowRegret <- plot_LowRegretStatistics(LowRegret_SummaryAll, col_name = "mean_log_climate_warming", y_axis = expression('log Δ'^"o"*'C yr'^"-1"*'')) + scale_y_reverse()

ggComparison_OceanAcidification_LowRegret <- plot_LowRegretStatistics(LowRegret_SummaryAll, col_name = "mean_ocean_acidification", y_axis = expression('Δ pH yr'^"-1"*'')) + scale_y_reverse()
ggComparison_LogOceanAcidification_LowRegret <- plot_LowRegretStatistics(LowRegret_SummaryAll, col_name = "mean_log_ocean_acidification", y_axis = expression('log Δ pH yr'^"-1"*'')) + scale_y_reverse()

ggComparison_OxygenDecline_LowRegret <- plot_LowRegretStatistics(LowRegret_SummaryAll, col_name = "mean_oxygen_decline", y_axis = expression('Δ mol m'^"-3"*' yr'^"-1"*'')) + scale_y_reverse()
ggComparison_LogOxygenDecline_LowRegret <- plot_LowRegretStatistics(LowRegret_SummaryAll, col_name = "mean_log_oxygen_decline", y_axis = expression('Δ mol m'^"-3"*' yr'^"-1"*'')) + scale_y_reverse()

ggComparison_ClimateVelocity_LowRegret <- plot_LowRegretStatistics(LowRegret_SummaryAll, col_name = "median_velocity", y_axis = expression('km yr'^"-1"*''))
ggComparison_LogClimateVelocity_LowRegret <- plot_LowRegretStatistics(LowRegret_SummaryAll, col_name = "mean_log_velocity", y_axis = expression('log km yr'^"-1"*''))
```
Cost and area:
``` {r}
ggComparison_Cost_LowRegret + ggComparison_Area_LowRegret + plot_layout(guides = "collect")
```
Climate warming:
``` {r}
ggComparison_Warming_LowRegret + ggComparison_LogWarming_LowRegret + plot_layout(guides = "collect")
```
Ocean acidification:
```{r}
ggComparison_OceanAcidification_LowRegret + ggComparison_LogOceanAcidification_LowRegret + plot_layout(guides = "collect")
```
Declining oxygen concentration:
```{r}
ggComparison_OxygenDecline_LowRegret + ggComparison_LogOxygenDecline_LowRegret + plot_layout(guides = "collect")
```
Climate velocity:
```{r}
ggComparison_ClimateVelocity_LowRegret + ggComparison_LogClimateVelocity_LowRegret + plot_layout(guides = "collect")
```
Cohen's Kappa Correlation Matrix:
```{r include = FALSE}
list <- c("LowRegret_Feature_585", "LowRegret_Percentile_585", "LowRegret_Penalty_585")
object_list <- list() # empty list
solution_list <- list(LowRegret_Feature, LowRegret_Percentile, LowRegret_Penalty)
for (i in 1:length(list)) {
  obj <- select_solution(solution_list[[i]], list[i])
  object_list[[i]] <- obj
}
```
``` {r echo = FALSE, message = FALSE}
(matrix <- create_corrmatrix(object_list) %>% 
    plot_corrplot(., length(object_list)))
```

Then, I create low-regret areas across all three low-regret spatial designs... just for the kicks of it.
```{r include = FALSE}
LowRegret_sf <- LowRegretAll_sf %>% 
  dplyr::rename(solution_1 = SelectionAll) %>% 
  dplyr::mutate(solution_1 = as.logical(solution_1))
```
```{r echo = FALSE}
(ggLowRegret_All <- fSpatPlan_PlotSolution(LowRegret_sf, PUs, land) + ggtitle("Low-Regret Areas", subtitle = "Across all approaches"))
```
```{r include = FALSE}
df <- cbind(LowRegret_sf, cost)
summary_lr <- compute_summary(df, total_area, PU_size, run_name = "LowRegret_All", Cost = "Cost_squish")
```
```{r}
print(summary_lr)
```
